A Responsible Machine Learning Workflow\\{with Focus on Interpretable Models, Post-hoc Explanation, and Discrimination Testing

### 2020-02-28 article-information-2019.tex
final, final, final edits

% Authors, for the paper (add full first names)
\Author{\hl{Navdeep Gill} $^{1, \dagger}$, Patrick Hall $^{1,2,}$*$^{,\dagger}$, Kim Montgomery $^{1,\dagger}$ and Nicholas Schmidt $^{3,}$*$^{,\dagger}$} %%%Author order is not the same as on the system, please confirm which one is correct.
%Please carefully check the accuracy of names and affiliations.

@@

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad \hl{H2O.ai}, 2307 Leghorn St., Mountain View, CA, 94043, USA; navdeep.gill@h2o.ai (N.G.); kim.montgomery@h2o.ai (K.M.)\\%%%Please provide the department, university,zip code, city, country.
$^{2}$ \quad \hl{The George Washington University}, Department of Decision Sciences, 2201 G Street NW,
Washington, D.C., 20052, USA\\%%%Please provide the department, zip code, city, country.
$^{3}$ \quad \hl{BLDS, LLC}, 1608 Walnut St., Suite 1108, Philadelphia, PA, 19103, USA}%%%Please provide the department, university,zip code, city, country.

@@

Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that~perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds ({\hl{See}%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
%is the text color necessary? please confirm if need remove the link of the text,
%% we have removed the link of text, and keep web link
: 
% THESE HYPERLINKS ARE NECESSARY - PH
\href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}
{Responsible Artificial Intelligence}, 
\href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}
{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, 
\href{https://www.pwc.com/gx/en/issues/data-and-analytics/artificial-intelligence/what-is-responsible-ai.html}
{PwC's Responsible AI}, 
\href{https://ai.google/responsibilities/responsible-ai-practices/}
{Responsible AI Practices}). Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is becoming clear. Machine learning (ML) models, a common application of AI, can present serious risks. ML~models can be inaccurate and unappealable %INCORRECT - PH -- Please confirm that the intended meaning has been retained.
black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop} ({\hl{See}% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
: 
\href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}
{\textit{When a Computer Program Keeps You in Jail}}). ML models can perpetuate and exacerbate discrimination \cite{feldman2015certifying,dwork2012fairness,gender_shades}, and ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml,model_stealing,membership_inference,shokri2019privacy}. This~manuscript makes no claim that these interdependent issues of ML opaqueness, discrimination, privacy harms, and security vulnerabilities have been resolved, even as singular entities, and much less as complex intersectional phenomena. However, Sections \ref{sec:m_and_m}--\ref{sec:disc} do propose some specific technical countermeasures, mostly in the form of interpretable models, post-hoc explanation, and disparate impact (DI) and discrimination testing, that responsible practitioners can use to address a subset of these vexing problems ({\hl{This text and associated software are not, and should not be construed as, legal advice or regulatory compliance requirements}}), ({\hl{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve Supervision and Regulation (SR) Letter 11-7, the European Union (EU) General Data Protection Regulation (GDPR) Article} 22, or other laws and regulations \cite{ff_interpretability}.})%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this.

@@

The simulated data (see Section~\ref{ssec:sim_data}) are based on the well-known Friedman datasets. Its~known feature importance and augmented discrimination characteristics are used to gauge the validity of interpretable modeling, post-hoc explanation, and discrimination testing techniques \cite{friedman1979tree,friedman1991multivariate}. The~mortgage data (see Section~\ref{ssec:mort_data}) are sourced from the Home Mortgage Disclosure Act (HMDA) database, a fairly realistic data source for demonstrating the template workflow \cite{hmda} ({\hl{See:} 
\href{https://www.consumerfinance.gov/data-research/hmda/}
{Mortgage data (HMDA})}). %%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
To provide a sense of fit differences, performance is compared on simulated data and collected mortgage data between the more interpretable constrained ML models and the less interpretable unconstrained ML models. Because the unconstrained ML models, gradient boosting machines (GBMs, e.g., \cite{friedman2001greedy,esl}) and artificial neural networks (ANNs, e.g., \cite{recht2011hogwild,hinton2012improving,sutskever2013importance,zeiler2012adadelta}), do not exhibit convincing accuracy benefits on the simulated or mortgage data and can also present the unmitigated risks discussed above, further explanation and discrimination analyses are applied only to the constrained, interpretable ML models \cite{please_stop,fair_washing,scaffolding}. Here, monotonic gradient boosting machines ({\hl{as implemented in} 
%\href{https://xgboost.readthedocs.io/en/latest/tutorials/monotonic.html}
{\texttt{\hl{XGBoost}}} or 
%\href{https://github.com/h2oai/h2o-3/blob/master/h2o-py/demos/H2O_tutorial_gbm_monotonicity.ipynb}
{\texttt{\hl{h2o}}}.}) %% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. % we have removed the link of text, and keep web link
(MGBMs, see Section~\ref{ssec:mgbm}) and explainable neural networks (XNNs, see Section~\ref{ssec:xnn} \cite{,wf_xnn,yang2019enhancing}) will serve as those more interpretable models for subsequent explanatory and discrimination analyses. MGBM and XNN interpretable model architectures are selected for the example workflow because they are straightforward variants of popular unconstrained ML models. If~practitioners are working with GBM and ANN models, it should be relatively uncomplicated to also evaluate the constrained versions of these models.

The same can be said of the selected explanation methods and discrimination tests. Due to their post-hoc nature, they can often be shoe-horned into existing ML workflows and pipelines. Presented explanation techniques include partial dependence (PD) and individual conditional expectation (ICE) (see Section~\ref{ssec:pd_ice}) and Shapley values (see Section~\ref{ssec:shap}) \cite{esl,ice_plots,shapley,tree_shap}. PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models. Discrimination testing methods discussed (see Section~\ref{ssec:di}) include adverse impact ratio (AIR), marginal effect (ME), and standardized mean difference (SMD)  \cite{feldman2015certifying,cohen1988statistical,cohen1992power}. ({\hl{See: Part 1607 - Uniform Guidelines on Employee Selection Procedures} (1978) \textsection 1607.4.}).
%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
 Accuracy and other confusion matrix measures are also reported by demographic segment \cite{zafar2017fairness}. All outlined materials and methods are implemented in open source Python code, and are made available on GitHub (see Section~\ref{ssec:soft}).

@@

Shapley explanations are a class of additive, locally accurate feature contribution measures with long-standing theoretical support \cite{shapley,shapley1988shapley}. Shapley explanations are the only known locally accurate and globally consistent feature contribution values, meaning that Shapley explanation values for input features always sum to the model's prediction, $g(\mathbf{x})$, for any instance $\mathbf{x}$, and that Shapley explanation values should not decrease in magnitude for some instance of $x_j$ when $g$ is changed such that $x_j$ truly makes a stronger contribution to $g(\mathbf{x})$ \cite{shapley,tree_shap}. Shapley values can be estimated in different ways, many of which are intractable for datasets with large numbers of input features. Tree~Shapley Additive Explanations (SHAP) is a specific implementation of Shapley explanations that relies on traversing internal decision tree structures to efficiently estimate the contribution of each $x_j$ for some $g(\mathbf{x})$ \cite{tree_shap}. Tree SHAP has been implemented in popular gradient boosting libraries such as 
%\href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}
{\texttt{\hl{h2o}}}, 
%\href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}
{\texttt{\hl{LightGBM}}}, and 
%\href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}
{\texttt{\hl{XGBoost}}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in Section~\ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}. Deep SHAP is an approximate Shapley value technique that creates SHAP values for ANNs \cite{shapley}. Deep SHAP is implemented in the 
%\href{https://github.com/slundberg/shap}
{\texttt{\hl{shap}}} package and is used to generate SHAP values for the two $g^{XNN}$ models discussed in Section~\ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}. For more information pertaining to the calculation of Shapley values, see Appendices \ref{a_ssec:not} and \ref{a_ssec:shap}.%% we have removed the link of text, and keep web link

@@

Because many current technical discussions of fairness in ML appear inconclusive ({See:} 
\href{https://www.youtube.com/watch?v=jIXIuYdnyyk}
{Tutorial: 21 Fairness Definitions and Their Politics}.)%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
, this text will draw on regulatory and legal standards that have been used for years in regulated, high-stakes employment and financial decisions. The~discussed measures are also representative of fair lending analyses and pair well with the mortgage data. (See~Appendix \ref{a_sec:types} for a brief discussion regarding different types of discrimination in US legal and regulatory settings, and Appendix \ref{a_sec:prac_v_stat} for remarks on practical vs. statistical significance for discrimination measures.) One such common measure of DI used in US litigation and regulatory settings is ME. ME is simply the difference between the percent of the control group members receiving a favorable outcome and the percent of the protected class members receiving a favorable outcome: 
\begin{equation}
\label{eq:me}
\begin{aligned}
\text{ME} \equiv 100 \cdot (\text{Pr}(\hat{\mathbf{y}} = 1| X_c = 1)  - \text{Pr}(\hat{\mathbf{y}} = 1 | X_p = 1))
\end{aligned}
\end{equation}
 where $\hat{\mathbf{y}}$ are the model decisions, $X_p$ and $X_c$ represent binary markers created from some demographic attribute, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and $\text{Pr}(\cdot)$ is the operator for conditional probability.  ME is a favored DI measure used by the US Consumer Financial Protection Bureau (CFPB), the primary agency charged with regulating fair lending laws at the largest US lending institutions and for various other participants in the consumer financial market. ({See: 
\href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}
{Supervisory Highlights, Issue 9, Fall 2015}.) %%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
 Another important DI measure is AIR, more~commonly known as a {relative risk ratio} in settings outside of regulatory compliance.

\begin{equation}
\label{eq:air}
\begin{aligned}
\text{AIR} \equiv \frac{\text{Pr}(\hat{\mathbf{y}} = 1 | X_p = 1)}{\text{Pr}(\hat{\mathbf{y}} = 1| X_c = 1)}
\end{aligned}
\end{equation}
 AIR is equal to the ratio of the proportion of the protected class that receives a favorable outcome and the proportion of the control class that receives a favorable outcome. Statistically significant AIR values below 0.8 can be considered {prima facie} evidence of discrimination. An additional long-standing and pertinent measure of DI is SMD. SMD is often used to assess disparities in continuous features, such~as income differences in employment analyses, or interest rate differences in lending. It originates from work on statistical power, and is more formally known as {Cohen’s d}. SMD is equal to the difference in the average protected class outcome, $\bar{\hat{\mathbf{y}}}_p$, minus the control class outcome, $\bar{\hat{\mathbf{y}}}_c$, divided by a measure of the standard deviation of the population, $\sigma_{\hat{\mathbf{y}}}$. ({\hl{There are several measures of the standard deviation of the score that are typically used}%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
: \textbf{\hl{1}.} the standard deviation of the population, irrespective of protected class status, \textbf{\hl{2}.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{\hl{3}.} a pooled standard deviation, using the standard deviations for each of the two groups with weights.}) Cohen defined values of this measure to have {small}, {medium}, and {large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively. %is bold necessary?

@@

Python code to reproduce discussed results is available at: {\url{https://github.com/h2oai/article-information-2019}}. The primary Python packages employed are:  
%\href{https://numpy.org/}
{\texttt{\hl{numpy}}} 1.14.5 and 
%\href{https://pandas.pydata.org/}
{\texttt{\hl{pandas}}} 0.22.0 for data manipulation, 
%\href{https://github.com/h2oai/h2o-3}
{\texttt{\hl{h2o}}} 3.26.0.9, 
%\href{https://keras.io/}
{\texttt{\hl{Keras}}} 2.3.1, 
%\href{https://github.com/slundberg/shap}
{\texttt{\hl{shap}}} 0.31.0, and 
%\href{https://www.tensorflow.org/}
{\texttt{\hl{tensorflow}}} 1.14.0 for modeling, explanation, and discrimination testing, and typically 
%\href{https://matplotlib.org/}
{\texttt{\hl{matplotlib}}} 2.2.2 for plotting.

@@

\begin{table}[H]%% we have removed the link of text
	\caption{Fit measures for $g^\text{GBM}(\mathbf{X})$, $g^\text{MGBM}(\mathbf{X})$, $g^\text{ANN}(\mathbf{X})$, and $g^\text{XNN}(\mathbf{X})$ on the simulated test data. Arrows indicate the direction of improvement for each measure and the best result in each column is displayed in bold font.}
	\centering
	\begin{tabular}{ccccccccc}
		\toprule
		\textbf{Model} & 
		%\href{https://scikit-learn.org/stable/modules/model_evaluation.html#accuracy-score}
		{\textbf{\hl{Accuracy}}}~$\uparrow$ & 
		%\href{https://en.wikipedia.org/wiki/Receiver_operating_characteristic#Area_under_the_curve}
		{\textbf{\hl{AUC}}}~$\uparrow$ & 
		%\href{https://en.wikipedia.org/wiki/F1_score}
		{\textbf{\hl{F1}}}~$\uparrow$ & 
		%\href{https://scikit-learn.org/stable/modules/model_evaluation.html#log-loss}
		{\textbf{\hl{Logloss}}}~$\downarrow$ & 
		%\href{https://en.wikipedia.org/wiki/Matthews_correlation_coefficient}
		{\textbf{\hl{MCC}}}~$\uparrow$ & 
		%\href{https://en.wikipedia.org/wiki/Root-mean-square_deviation}
		{\textbf{\hl{RMSE}}}~$\downarrow$ & 
		%\href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}
		{\textbf{\hl{Sensitivity}}}~$\uparrow$ & 
		%\href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}
		{\textbf{\hl{Specificity}}}~$\uparrow$ \\
		\midrule
		$g^\text{GBM}$		& 0.757 & 0.847 & 0.779 & 0.486 & 0.525 & 0.400 & 0.858 & \textbf{0.657} \\
		$g^\text{MGBM}$		& 0.744 & 0.842 & 0.771 & 0.502 & 0.504 & 0.407 & 0.864 & 0.625 \\
		$g^\text{ANN}$		& 0.757	& 0.850	& 0.779 & 0.480 & 0.525 & 0.398 & 0.858 & \textbf{0.657} \\
		$g^\text{XNN}$		& \textbf{0.758}	& \textbf{0.851}	& \textbf{0.781} & \textbf{0.479} & \textbf{0.528} & \textbf{0.397} & \textbf{0.867} & 0.648 \\
		\bottomrule
		\label{tab:sim_acc}
	\end{tabular}
\end{table}

@@

\begin{figure}[H]
\centering
	%\includegraphics[width=15cm]{img/sim_xnn_glob_ridge.png}
	\includegraphics[width=17cm]{img/sim_xnn_glob_ridge.png}
	\caption{Output layer $\gamma_k$ weights, corresponding $n_k$ subnetwork ridge functions, and associated projection layer $\beta_{k,j}$ weights for $g^\text{XNN}$ on the simulated data.}
	\label{fig:sim_xnn_glob_ridge}
\end{figure}

@@

\begin{figure}[H]
\includegraphics[width=17cm]{img/mort_xnn_glob_ridge.png}
\caption{Output layer $\gamma_k$ weights, corresponding $n_k$ ridge functions, and associated projection layer $\beta_{k,j}$ weights for $g^\text{XNN}$ on the mortgage data.}
\label{fig:mort_xnn_glob_ridge}
\end{figure}

>>>>

% Authors, for the paper (add full first names)
\Author{{Navdeep Gill} $^{1, \dagger}$, Patrick Hall $^{1,2,}$*$^{,\dagger}$, Kim Montgomery $^{1,\dagger}$ and Nicholas Schmidt $^{3,}$*$^{,\dagger}$} %%%Author order is not the same as on the system, please confirm which one is correct.
%Please carefully check the accuracy of names and affiliations.

@@

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad {H2O.ai}, Mountain View 94043, CA, USA; navdeep.gill@h2o.ai (N.G.); kim.montgomery@h2o.ai (K.M.)\\
$^{2}$ \quad {Department of Decision Sciences, The George Washington University}, Washington 20052, DC,  USA\\%%%Please provide the department, zip code, city, country.
$^{3}$ \quad {BLDS, LLC}, Philadelphia 19103, PA, USA}

@@

Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that~perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds (e.g.,
% THESE HYPERLINKS ARE NECESSARY - PH
\href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}
{Responsible Artificial Intelligence}, 
\href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}
{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, 
\href{https://www.pwc.com/gx/en/issues/data-and-analytics/artificial-intelligence/what-is-responsible-ai.html}
{PwC's Responsible AI}, 
\href{https://ai.google/responsibilities/responsible-ai-practices/}
{Responsible AI Practices}). 
Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is becoming clear. Machine learning (ML) models, a common application of AI, can present serious risks. ML~models can be inaccurate and unappealable %INCORRECT - PH -- Please confirm that the intended meaning has been retained.
black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop} (e.g.,
\href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}
{\textit{When a Computer Program Keeps You in Jail}}). ML models can perpetuate and exacerbate discrimination \cite{feldman2015certifying,dwork2012fairness,gender_shades}, and ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml,model_stealing,membership_inference,shokri2019privacy}. This~manuscript makes no claim that these interdependent issues of ML opaqueness, discrimination, privacy harms, and security vulnerabilities have been resolved, even as singular entities, and much less as complex intersectional phenomena. However, Sections \ref{sec:m_and_m}--\ref{sec:disc} do propose some specific technical countermeasures, mostly in the form of interpretable models, post-hoc explanation, and disparate impact (DI) and discrimination testing, that responsible practitioners can use to address a subset of these vexing problems. (In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may also be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve Supervision and Regulation (SR) Letter 11-7, the European Union (EU) General Data Protection Regulation (GDPR) Article 22, or other laws and regulations \cite{ff_interpretability}. Note that this text and associated software are not, and should not be construed as, legal advice or regulatory compliance requirements.)

@@

The simulated data (see Section~\ref{ssec:sim_data}) are based on the well-known Friedman datasets. Its~known feature importance and augmented discrimination characteristics are used to gauge the validity of interpretable modeling, post-hoc explanation, and discrimination testing techniques \cite{friedman1979tree,friedman1991multivariate}. The~mortgage data (see Section~\ref{ssec:mort_data}) are sourced from the Home Mortgage Disclosure Act (HMDA) database, a fairly realistic data source for demonstrating the template workflow \cite{hmda} (see
\href{https://www.consumerfinance.gov/data-research/hmda/}
{Mortgage data (HMDA)}). %%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
To provide a sense of fit differences, performance is compared on simulated data and collected mortgage data between the more interpretable constrained ML models and the less interpretable unconstrained ML models. Because the unconstrained ML models, gradient boosting machines (GBMs, e.g., \cite{friedman2001greedy,esl}) and artificial neural networks (ANNs, e.g., \cite{recht2011hogwild,hinton2012improving,sutskever2013importance,zeiler2012adadelta}), do not exhibit convincing accuracy benefits on the simulated or mortgage data and can also present the unmitigated risks discussed above, further explanation and discrimination analyses are applied only to the constrained, interpretable ML models \cite{please_stop,fair_washing,scaffolding}. Here, monotonic gradient boosting machines 
(MGBMs, as implemented in 
%\href{https://xgboost.readthedocs.io/en/latest/tutorials/monotonic.html}
\texttt{XGBoost} or 
%\href{https://github.com/h2oai/h2o-3/blob/master/h2o-py/demos/H2O_tutorial_gbm_monotonicity.ipynb}
\texttt{h2o}, see Section~\ref{ssec:mgbm}) 
and explainable neural networks (XNNs, e.g., \cite{,wf_xnn,yang2019enhancing}, see Section~\ref{ssec:xnn}) will serve as those more interpretable models for subsequent explanatory and discrimination analyses. MGBM and XNN interpretable model architectures are selected for the example workflow because they are straightforward variants of popular unconstrained ML models. If~practitioners are working with GBM and ANN models, it should be relatively uncomplicated to also evaluate the constrained versions of these models.

The same can be said of the selected explanation methods and discrimination tests. Due to their post-hoc nature, they can often be shoe-horned into existing ML workflows and pipelines. Presented explanation techniques include partial dependence (PD) and individual conditional expectation (ICE) (see Section~\ref{ssec:pd_ice}) and Shapley values (see Section~\ref{ssec:shap}) \cite{esl,ice_plots,shapley,tree_shap}. PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models. Discrimination testing methods discussed (see Section~\ref{ssec:di}) include adverse impact ratio
(AIR, see Part 1607---Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4), 
marginal effect (ME), and standardized mean difference (SMD)  \cite{feldman2015certifying,cohen1988statistical,cohen1992power}.
%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
Accuracy and other confusion matrix measures are also reported by demographic segment \cite{zafar2017fairness}. All outlined materials and methods are implemented in open source Python code, and are made available on GitHub (see Section~\ref{ssec:soft}).

@@

Shapley explanations are a class of additive, locally accurate feature contribution measures with long-standing theoretical support \cite{shapley,shapley1988shapley}. Shapley explanations are the only known locally accurate and globally consistent feature contribution values, meaning that Shapley explanation values for input features always sum to the model's prediction, $g(\mathbf{x})$, for any instance $\mathbf{x}$, and that Shapley explanation values should not decrease in magnitude for some instance of $x_j$ when $g$ is changed such that $x_j$ truly makes a stronger contribution to $g(\mathbf{x})$ \cite{shapley,tree_shap}. Shapley values can be estimated in different ways, many of which are intractable for datasets with large numbers of input features. Tree~Shapley Additive Explanations (SHAP) is a specific implementation of Shapley explanations that relies on traversing internal decision tree structures to efficiently estimate the contribution of each $x_j$ for some $g(\mathbf{x})$ \cite{tree_shap}. Tree SHAP has been implemented in popular gradient boosting libraries such as 
%\href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}
{\texttt{{h2o}}}, 
%\href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}
{\texttt{{LightGBM}}}, and 
%\href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}
{\texttt{{XGBoost}}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in Section~\ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}. Deep SHAP is an approximate Shapley value technique that creates SHAP values for ANNs \cite{shapley}. Deep SHAP is implemented in the 
%\href{https://github.com/slundberg/shap}
{\texttt{{shap}}} package and is used to generate SHAP values for the two $g^{XNN}$ models discussed in Section~\ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}. For more information pertaining to the calculation of Shapley values, see Appendices \ref{a_ssec:not} and \ref{a_ssec:shap}.%% we have removed the link of text, and keep web link

@@

Because many current technical discussions of fairness in ML appear inconclusive (e.g., 
\href{https://www.youtube.com/watch?v=jIXIuYdnyyk}{Tutorial: 21 Fairness Definitions and Their Politics}), this text will draw on regulatory and legal standards that have been used for years in regulated, high-stakes employment and financial decisions. The~discussed measures are also representative of fair lending analyses and pair well with the mortgage data. (See~Appendix \ref{a_sec:types} for a brief discussion regarding different types of discrimination in US legal and regulatory settings, and Appendix \ref{a_sec:prac_v_stat} for remarks on practical vs. statistical significance for discrimination measures.) One such common measure of DI used in US litigation and regulatory settings is ME. ME is simply the difference between the percent of the control group members receiving a favorable outcome and the percent of the protected class members receiving a favorable outcome: 
\begin{equation}
\label{eq:me}
\begin{aligned}
\text{ME} \equiv 100 \cdot (\text{Pr}(\hat{\mathbf{y}} = 1| X_c = 1)  - \text{Pr}(\hat{\mathbf{y}} = 1 | X_p = 1))
\end{aligned}
\end{equation}
 where $\hat{\mathbf{y}}$ are the model decisions, $X_p$ and $X_c$ represent binary markers created from some demographic attribute, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and $\text{Pr}(\cdot)$ is the operator for conditional probability.  ME is a favored DI measure used by the US Consumer Financial Protection Bureau (CFPB), the primary agency charged with regulating fair lending laws at the largest US lending institutions and for various other participants in the consumer financial market (see  
\href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}
{\textit{Supervisory Highlights}, Issue 9, Fall 2015}).
 Another important DI measure is AIR, more~commonly known as a relative risk ratio in settings outside of regulatory compliance.

\begin{equation}
\label{eq:air}
\begin{aligned}
\text{AIR} \equiv \frac{\text{Pr}(\hat{\mathbf{y}} = 1 | X_p = 1)}{\text{Pr}(\hat{\mathbf{y}} = 1| X_c = 1)}
\end{aligned}
\end{equation}
 AIR is equal to the ratio of the proportion of the protected class that receives a favorable outcome and the proportion of the control class that receives a favorable outcome. Statistically significant AIR values below 0.8 can be considered {prima facie} evidence of discrimination. An additional long-standing and pertinent measure of DI is SMD. SMD is often used to assess disparities in continuous features, such~as income differences in employment analyses, or interest rate differences in lending. It originates from work on statistical power, and is more formally known as {Cohen’s d}. SMD is equal to the difference in the average protected class outcome, $\bar{\hat{\mathbf{y}}}_p$, minus the control class outcome, $\bar{\hat{\mathbf{y}}}_c$, divided by a measure of the standard deviation of the population, $\sigma_{\hat{\mathbf{y}}}$. ({{There are several measures of the standard deviation of the score that are typically used}%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
: \textbf{{1}.} the standard deviation of the population, irrespective of protected class status, \textbf{{2}.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{{3}.} a pooled standard deviation, using the standard deviations for each of the two groups with weights.}) Cohen defined values of this measure to have {small}, {medium}, and {large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively. %is bold necessary?

@@

Python code to reproduce discussed results is available at: {\url{https://github.com/h2oai/article-information-2019}}. The primary Python packages employed are:  
%\href{https://numpy.org/}
{\texttt{{numpy}}} 1.14.5 and 
%\href{https://pandas.pydata.org/}
{\texttt{{pandas}}} 0.22.0 for data manipulation, 
%\href{https://github.com/h2oai/h2o-3}
{\texttt{{h2o}}} 3.26.0.9, 
%\href{https://keras.io/}
{\texttt{{Keras}}} 2.3.1, 
%\href{https://github.com/slundberg/shap}
{\texttt{{shap}}} 0.31.0, and 
%\href{https://www.tensorflow.org/}
{\texttt{{tensorflow}}} 1.14.0 for modeling, explanation, and discrimination testing, and typically 
%\href{https://matplotlib.org/}
{\texttt{{matplotlib}}} 2.2.2 for plotting.

@@

\begin{table}[H]%% we have removed the link of text
	\caption{Fit measures for $g^\text{GBM}(\mathbf{X})$, $g^\text{MGBM}(\mathbf{X})$, $g^\text{ANN}(\mathbf{X})$, and $g^\text{XNN}(\mathbf{X})$ on the simulated test data. Arrows indicate the direction of improvement for each measure and the best result in each column is displayed in bold font.}
	\centering
	\begin{tabular}{ccccccccc}
		\toprule
		\textbf{Model} & 
		%\href{https://scikit-learn.org/stable/modules/model_evaluation.html#accuracy-score}
		{\textbf{{Accuracy}}}~$\uparrow$ & 
		%\href{https://en.wikipedia.org/wiki/Receiver_operating_characteristic#Area_under_the_curve}
		{\textbf{{AUC}}}~$\uparrow$ & 
		%\href{https://en.wikipedia.org/wiki/F1_score}
		{\textbf{{F1}}}~$\uparrow$ & 
		%\href{https://scikit-learn.org/stable/modules/model_evaluation.html#log-loss}
		{\textbf{{Logloss}}}~$\downarrow$ & 
		%\href{https://en.wikipedia.org/wiki/Matthews_correlation_coefficient}
		{\textbf{{MCC}}}~$\uparrow$ & 
		%\href{https://en.wikipedia.org/wiki/Root-mean-square_deviation}
		{\textbf{{RMSE}}}~$\downarrow$ & 
		%\href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}
		{\textbf{{Sensitivity}}}~$\uparrow$ & 
		%\href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}
		{\textbf{{Specificity}}}~$\uparrow$ \\
		\midrule
		$g^\text{GBM}$		& 0.757 & 0.847 & 0.779 & 0.486 & 0.525 & 0.400 & 0.858 & \textbf{0.657} \\
		$g^\text{MGBM}$		& 0.744 & 0.842 & 0.771 & 0.502 & 0.504 & 0.407 & 0.864 & 0.625 \\
		$g^\text{ANN}$		& 0.757	& 0.850	& 0.779 & 0.480 & 0.525 & 0.398 & 0.858 & \textbf{0.657} \\
		$g^\text{XNN}$		& \textbf{0.758}	& \textbf{0.851}	& \textbf{0.781} & \textbf{0.479} & \textbf{0.528} & \textbf{0.397} & \textbf{0.867} & 0.648 \\
		\bottomrule
		\label{tab:sim_acc}
	\end{tabular}
\end{table}

@@

% start a new page
\newpage
% change it to landscape
\paperwidth=\pdfpageheight
\paperheight=\pdfpagewidth
\pdfpageheight=\paperheight
\pdfpagewidth=\paperwidth
\newgeometry{layoutwidth=297mm,layoutheight=210 mm, left=2.7cm,right=2.7cm,top=1.8cm,bottom=1.5cm, includehead,includefoot}
\fancyheadoffset[LO,RE]{0cm}
\fancyheadoffset[RO,LE]{0cm}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[H]
\centering
	%\includegraphics[width=15cm]{img/sim_xnn_glob_ridge.png}
	\includegraphics[width=23cm]{img/sim_xnn_glob_ridge.png}
	\caption{Output layer $\gamma_k$ weights, corresponding $n_k$ subnetwork ridge functions, and associated projection layer $\beta_{k,j}$ weights for $g^\text{XNN}$ on the simulated data.}
	\label{fig:sim_xnn_glob_ridge}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% change everything back
\newpage
\restoregeometry
\paperwidth=\pdfpageheight
\paperheight=\pdfpagewidth
\pdfpageheight=\paperheight
\pdfpagewidth=\paperwidth
\headwidth=\textwidth

@@

% start a new page
\newpage
% change it to landscape
\paperwidth=\pdfpageheight
\paperheight=\pdfpagewidth
\pdfpageheight=\paperheight
\pdfpagewidth=\paperwidth
\newgeometry{layoutwidth=297mm,layoutheight=210 mm, left=2.7cm,right=2.7cm,top=1.8cm,bottom=1.5cm, includehead,includefoot}
\fancyheadoffset[LO,RE]{0cm}
\fancyheadoffset[RO,LE]{0cm}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
